---
title: "Example"
author: "Michael Li"
date: "January 10, 2017"
output: html_document
---

### Simulating Animal movement using depmixS4

DepmixS4 is one of many package to fit hidden Markov models (HMM). In this example, we will demostrate how to simulate and fit covariate-dependent transition HMM using depmixS4.

```{r setup}
tempdat <- data.frame(time=rep(1:24,500),y=NA)
head(tempdat)
```

Create a depmix object

```{r temp depmix object, echo=FALSE}
library(depmixS4)
tempmod <- (depmix(y~1  # response 
  , nstates = 2 # number of states
  , transition=~cos(2*pi*time/24)+sin(2*pi*time/24) # transition structure 
  , family= gaussian()
  , data=tempdat
)
)

summary(tempmod)
```

Set HMM parameters 

```{r set parameters}
setpars(tempmod, c(0.1, 0.9 # Initial prob parameters
  , 0 , 1 , 0 , 1 , 0 , 1 , 0 , -1 , 0 , -1 , 0 , -1 # Sinusoidal parameters for state switch (mlogit)
  , 0 , 1 , 2 , 1 # emission distribution parameters 
  )
)

summary(tempmod)
```

Simulate hidden states and response data and put it back into a data frame. Strongly recommend simulating the response data using R, especially for user defined emission distrubtions.

```{r simulate hmm}
source("simulate.R")
simdat <- simhmm(tempmod,nsim=1)
simdf <- data.frame(y=simdat@response[[1]][[1]]@y
  , states = simdat@states
  , time = tempdat$time)

head(simdf)
```

### Fitting HMM

Create a new depmix object and fit the simulated data above 

```{r fit depmix}
source("simulate.R")
simdf <- data.frame(y=simdat@response[[1]][[1]]@y
  , states = simdat@states
  , time = tempdat$time)

mod <- (depmix(y~1  # response 
  , nstates = 2 # number of states
  , transition=~cos(2*pi*time/24)+sin(2*pi*time/24) # transition structure 
  , family= gaussian()
  , data=simdf
)
)

fitmod <- fit(mod,verbose=FALSE)
summary(fitmod)
```
